
%\addcontentsline{toc}{chapter}{Acknowledgement}

\begin{acknowledgement}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)

I want to extend my deepest gratitude to...

\begin{itemize}
\item Carl Thomé and Carlos Lordeloª, whose expertise in MIR, ML, and DSP has been pivotal to the progress and success of my thesis. Their invaluable guidance and unwavering encouragement have empowered me to overcome the challenges encountered throughout this journey. Their dedication and commitment to mentorship are truly inspiring, and their influence has left an indelible mark on my academic path. The opportunity to work under their mentorship has been a privilege and an honor. Their profound impact is beyond what words can encapsulate.

\vspace*{3mm}
\item Cody Hesse and Sebastian Löf

\vspace*{3mm}
\item Epidemic Sound

\vspace*{3mm}
\item The Music Technology Group (MTG) and Universitat Pompeu Fabra (UPF). My profound thanks go to Xavier Serra for his faith in accepting a non-traditional profile to the master's program. Additionally, I appreciate Pedro Ramoneda and Genís Plaja for their patient assistance throughout my journey.

\vspace*{3mm}
\item My colleague at My Sheet Music Transcriptions, Oriol López, for his exceptional support and flexibility throughout my thesis. His understanding and willingness to accommodate my academic commitments effectively allowed me to balance my work and research responsibilities.

\vspace*{3mm}
Thank you all for being excellent advisors and playing a crucial role in completing my thesis. I sincerely appreciate everything you have done for me and will always cherish the knowledge and experience I have gained under your guidance.

\vspace*{3mm}
\end{itemize}

\newpage
\end{acknowledgement}